---
title: "Shmistogram plots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Mixed data types

A shmistogram highlights point masses that may occur in a continuous variable, explicitly decomposing the distribution into a mixture of a multinomial and a piecewise uniform distribution. Such point masses are quite common:
- inconsistent rounding any continuous variable can introduce a mixture of point masses and relatively continuous observations
- "age of earning first driver's license" plausibly has structural modes at the legal minimum (which may vary by state) and otherwise vary continuously

Let's simulate a uniform distribution mixed with a two point masses and a couple missing observations:

```{r fig.width=9, fig.height=5}
set.seed(1)
unif = runif(n=100, min=-10, max=100)
multi = c(rep(0, 20), rep(42, 10), rep(NA, 2))
data = c(unif, multi)
par(mfrow = c(1, 2))
hist(data, border=NA, col='grey', main = 'Histogram')
shmistogram::shmistogram(data)
```

The histogram obscures the point masses somewhat. By contrast, the shmistogram highlights the modes using red line segments, and also visualizes the fraction of
the data that is missing.

### Variable bin width

The shmistogram's agglomerative binning produces a parsimonious density estimate with adaptive bin width:

```{r fig.width=9, fig.height=5}
set.seed(0)
data = c(
    rnorm(2000, mean=0, sd=10),
    rnorm(500, mean = 4, sd = 0.5)
)
par(mfrow = c(1, 2))
hist(data, 50, border=NA, col='grey', main = 'Histogram')
shmistogram::shmistogram(data)
```
